---
title: "Week1, Day1 Homework Joins and Tidyr"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(skimr)
```

1.1 Joins

Q1
```{r}
card_base <- read_csv("data/CardBase.csv") %>% clean_names
# neater way to do this

customer_base <- read_csv("data/CustomerBase.csv")
customer_base <- clean_names(customer_base)

fraud_base <- read_csv("data/FraudBase.csv")
fraud_base <- clean_names(fraud_base)

transaction_base <- read_csv("data/TransactionBase.csv")
transaction_base <- clean_names(transaction_base)
```

Q2
```{r}
left_join(card_base, customer_base, by="cust_id")
```
Q3    
```{r}
fraud_base
transaction_base
full_join(transaction_base, fraud_base,  by = "transaction_id")
```
Final row count shows that there were no frauds that didnt have a matching a transaction

Q4
```{r}
skim(transaction_base) 
```
```{r}
skim(card_base) 
```
```{r}
transaction_base %>% 
  select(credit_card_id)
```
#check above shows "credit_card_id" does exist ??

```{r}
  semi_join(card_base, transaction_base, by = c("card_number", "credit_card_id"))
  #semi_join(transaction_base,card_base, by=c("credit_card_id", "card_number"))
```

1.2 tidyr

Q5

```{r}
hat_obs <- read_csv("data/hat_observations.csv")
hat_obs
```

```{r}
new_hat_obs <- hat_obs %>% 
 separate(observation, c("hat_colour","hat_type"), sep = ",")
```

Q6
```{r}
final_hat_obs <- new_hat_obs %>% 
  unite(date, c("year", "month", "day"), sep="-") 
```

```{r}
final_hat_obs %>% 
    filter(hat_type == "beret") %>% 
    filter(observation_count == max(observation_count))
```
Note that homework answer interprets question slightly differently
hat_observations_date <- hat_observations %>% 
  unite("date", day:year, sep = "-") 

hat_observations_date %>% 
  filter(hat_type == "beret") %>% 
  group_by(date) %>% 
  summarise(total = sum(observation_count)) %>% 
  slice_max(total)

2. Extension

Q2.1
```{r}
card_base
customer_base
transaction_base
fraud_base
```
```{r}
left_join(card_base, customer_base, by="cust_id") %>% 
  left_join(transaction_base, by = c("card_number" = "credit_card_id")) %>% 
   right_join(fraud_base, by = "transaction_id")
  
```


Q2.2
```{r}
exam_scores <- read_csv("data/exam_scores.csv")
exam_scores
```

```{r}
exam_scores_longer <- exam_scores %>% 
 pivot_longer(cols = contains("Q"), names_to = "exam_question", values_to = "score")
```

```{r}
exam_scores_longer %>% 
  separate(exam_question, c("prefix", "question_num"), sep ="Q") %>% 
  select(-prefix)
```

