---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(janitor)
library(skimr)
```

# defining a 'mode' function 
```{r}
get_mode <- function(data){
  tabled_data <- table(data)
  table_names <- names(tabled_data)
  return( table_names[tabled_data == max(tabled_data)] )
}
```

```{r}
mobile_data <- read_csv("data/20190928-items.csv") %>% 
clean_names()
```

#explore dataset
```{r}
nrow(mobile_data)
ncol(mobile_data)
skim(mobile_data)
```

#Q2 find the brand with the highest number of phones in this dataset

# do this by arranging count of brand by order
```{r}
mobile_data %>% 
  group_by(brand) %>% 
  summarise(count_brand = n()) %>% 
  arrange(desc(count_brand)) %>% 
  head(10)
```
#do this by looking at the distribution
```{r}
mobile_data %>% 
  group_by(brand) %>% 
  summarise(count_brand = n()) %>% 
  ggplot(aes(x= brand, y = count_brand)) +
  geom_col() +
  coord_flip()
```
# do this using the mode function
```{r}
top_brand <- mobile_data %>% 
  summarise(mode = get_mode(brand)) %>% 
  pull()
```

# Q3 For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?

```{r}
stats_top_brand <- mobile_data %>% 
  filter(brand == top_brand) %>% 
  select(rating) %>% 
  summarise(r_mean = mean(rating),
            r_sd = sd(rating),
            r_max = max(rating),
            r_min = min(rating),
            r_count = n())
```

#what would a normal distribution look like?
```{r}
  #create a sequence covering range of the data
  x=seq(from = stats_top_brand$r_min,
        to = stats_top_brand$r_max,
        by = (stats_top_brand$r_max-stats_top_brand$r_min)/20
        )
 #create a normal distribution over this set of values using mean and sd from dataset  
 f1_x = dnorm(x = x, mean = stats_top_brand$r_mean, sd = stats_top_brand$r_sd) 
```
# put this data in a tibble to use later
```{r}
freq_dist <- tibble(
    x = x,
    f1_x = f1_x
  ) 
```


#plot this
```{r}
 ggplot()+
    aes(x = x, y = f1_x) +
 geom_line(col = "red")
 #geom_line(aes(x = x, y = f1_x),col = "red")
```



#plot a basic histogram. This is a frequency distribution
```{r}
mobile_data %>% 
  filter(brand == top_brand) %>% 
  select(rating) %>% 
  ggplot(aes(x=rating)) +
  geom_histogram(bins = 20)
```
#plot as a density plot using geom_density
```{r}
mobile_data %>% 
  filter(brand == top_brand) %>% 
  select(rating) %>% 
  ggplot() +
  geom_density(aes(x=rating)) +
  #overlay this with the freq distribution we calculated earlier
  geom_line(data = freq_dist, aes(x = x, y = f1_x),col = "red")
```

These two distributions look similar (ish). The data is more peaky than a normal distribution. Less spread than a normal distribution or variance about the mean is lower? 
Not sure how do we quantify the similarness?

#Q4 We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

```{r}
within_1sd <- mobile_data %>% 
  filter(brand == top_brand) %>% 
  #first normalise ratings
  mutate(norm_ratings = (rating-stats_top_brand$r_mean)/stats_top_brand$r_sd) %>% 
  #then filter
  filter(between(norm_ratings, left = -1, right = 1)) %>% 
  summarise(count = n()/ sum(mobile_data$brand == top_brand))
```

```{r}
within_2sd <- mobile_data %>% 
  filter(brand == top_brand) %>% 
  #first normalise ratings
  mutate(norm_ratings = (rating-stats_top_brand$r_mean)/stats_top_brand$r_sd) %>% 
  #then filter
  filter(between(norm_ratings, left = -2, right = 2)) %>% 
  summarise(count = n()/ sum(mobile_data$brand == top_brand))
```

```{r}
within_3sd <- mobile_data %>% 
  filter(brand == top_brand) %>% 
  #first normalise ratings
  mutate(norm_ratings = (rating-stats_top_brand$r_mean)/stats_top_brand$r_sd) %>% 
  #then filter
  filter(between(norm_ratings, left = -3, right = 3)) %>% 
  summarise(count = n()/ sum(mobile_data$brand == top_brand))
```

```{r}
within_1sd #more in here)
pnorm(q=1)-pnorm(q=-1)
within_2sd
pnorm(q=2)-pnorm(q=-2)
within_3sd #less out here
pnorm(q=3)-pnorm(q=-3)
```
#Q5 Investigate the use of the qqnorm() plot in R to determine whether data is normally distributed. What is a ‘quantile-quantile plot’? What do we expect to see in a qqnorm() plot if data is nearly normally distributed?

```{r}
  # this creates a uniform distribution
  norm_points=qunif(ppoints(stats_top_brand$r_count))*stats_top_brand$r_max

  # this creates a normal distribution 
  # this is a set of values across the range of ratings - distributed around a mean of zero
  norm_points <- rnorm(stats_top_brand$r_count)
  
  #, probs=seq(stats_top_brand$r_min,stats_top_brand$r_max,0.01))


 #create a normal distribution over this set of values using mean and sd from dataset  
 ggplot() +
 aes(x = norm_points) +
 geom_histogram(bins=20)   
```

```{r}
  qqnorm(norm_points)
```


```{r}
top_brand_rating <- mobile_data %>% 
  filter(brand == top_brand) %>% 
  select(rating) 
```

```{r}
 ggplot() +
 aes(x = top_brand_rating$rating) +
 geom_histogram(bins=20)   
```

```{r}
  qqnorm(top_brand_rating$rating)
```
```{r}
  qqnorm(top_brand_rating$rating)
```
```{r}
qqplot(norm_points,top_brand_rating$rating)
```

