---
title: "Load data and test plots
output: html_notebook
---

# two datasets
# Note that I manually edited data to remove header lines and remove provisional data from year # 2022. I didnt want to spend time learning how to do this today.

# the file structure is like this
#   yyyy  mm   tmax    tmin      af    rain     sun
#              degC    degC    days      mm   hours

# also merrily ignoring data quality flags so removing * and #

```{r}
library(tidyverse)
library(janitor)
library(colorfindr)
```

```{r}
lerwick <- read.table(here::here("raw_data","lerwickdata2.txt"), header = TRUE)
paisley <- read.table(here::here("raw_data","paisleydata2.txt"), header = TRUE)
eastbourne <- read.table(here::here("raw_data","eastbournedata2.txt"), header = TRUE)
```

# cleaning script for met data

```{r}
lerwick <- clean_names(lerwick) 

lerwick_clean <- lerwick %>% 
  mutate(tmax = as.double(str_remove_all(tmax, "\\*|#"))) %>%   
  mutate(tmin = as.double(str_remove_all(tmin, "\\*|#"))) %>%  
  mutate(af = as.double(str_remove_all(af, "\\*|#"))) %>%  
  mutate(rain = as.double(str_remove_all(rain, "\\*|#"))) %>% 
  mutate(sun = as.double(str_remove_all(sun, "\\*|#")))  

  
paisley <- clean_names(paisley)
paisley_clean <- paisley %>% 
  mutate(tmax = as.double(str_remove_all(tmax, "\\*|#"))) %>%   
  mutate(tmin = as.double(str_remove_all(tmin, "\\*|#"))) %>%  
  mutate(af = as.double(str_remove_all(af, "\\*|#"))) %>%  
  mutate(rain = as.double(str_remove_all(rain, "\\*|#"))) %>% 
  mutate(sun = as.double(str_remove_all(sun, "\\*|#")))  


eastbourne <- clean_names(eastbourne)
eastbourne_clean <- eastbourne %>% 
  mutate(tmax = as.double(str_remove_all(tmax, "\\*|#"))) %>%   
  mutate(tmin = as.double(str_remove_all(tmin, "\\*|#"))) %>%  
  mutate(af = as.double(str_remove_all(af, "\\*|#"))) %>%  
  mutate(rain = as.double(str_remove_all(rain, "\\*|#"))) %>% 
  mutate(sun = as.double(str_remove_all(sun, "\\*|#")))  

names(eastbourne)  
names(lerwick)
names(paisley)
```
```{r}
glimpse(lerwick_clean)
glimpse(paisley_clean)
glimpse(eastbourne_clean)
```



```{r}
# calculate long term averages

lerwick_avg <- lerwick_clean %>% 
  filter(yyyy %in% c(1991:2020)) %>%
  group_by (mm) %>% 
  summarise(across(everything(),
                     ~{mean(.x, na.rm = TRUE)})) %>%
  select(-yyyy) %>% 
  mutate(location = "lerwick") 
```

```{r}
# calculate long term averages
paisley_avg <- paisley_clean %>%
  filter(yyyy %in% c(1991:2020)) %>%
  group_by (mm) %>% 
  summarise(across(everything(),
                     ~{mean(.x, na.rm = TRUE)})) %>%
  select(-yyyy) %>% 
  mutate(location = "paisley") 
```

```{r}
# calculate long term averages
eastbourne_avg <- eastbourne_clean %>%
  filter(yyyy %in% c(1991:2020)) %>%
  group_by (mm) %>% 
  summarise(across(everything(),
                     ~{mean(.x, na.rm = TRUE)})) %>%
  select(-yyyy) %>% 
  mutate(location = "eastbourne") 
```

```{r}
# merge them together
met_data <-  full_join(lerwick_avg,paisley_avg)
met_data <- full_join(met_data,eastbourne_avg)
met_data = pivot_longer(met_data, c(tmax,tmin,af,rain,sun), 
                        names_to = "param", 
                        values_to = "data") %>% 
  mutate(param = recode(param, "tmax" = "Mean Daily Max Temperature",
                        "tmin" = "Mean Daily Min Temperature",
                        "af"   = "Air Frost",
                        "rain" = "Rainfall",
                        "sun" = "Sunshine"))
```
```{r}
write_csv(met_data,here::here(
    "clean_data", "met_data.csv"))
```



```{r}
#create example plot

sel_param = "Sunshine"
sel_location = c("lerwick","eastbourne")

met_select <- met_data %>% 
  filter(location %in% sel_location) %>% 
  filter(param == sel_param)

  ggplot(met_select) +
    (aes(x = mm, y = data, colour = location)) +
  geom_line() +
  geom_point() +
  ylab(sel_param) +
  xlab("Month") +  
  scale_x_continuous(n.breaks = 12)  
```

# create theme colours

# extract palette from screenshot of minty theme
```{r}
# If you rerun this colour selection will change
#metpal <- get_colors(here::here("minty_screenshot.png"))%>%  make_palette(n = 10)

#colours <- metpal[c(7,8,9)]
#metpal2 <- as.data.frame(colours)
#write_csv(metpal2,here::here("met_palette.csv"))
```

