---
title: "Bootstrap Homework"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(infer)
library(janitor)
library(e1071) #for skewness
library(plotrix)
```

#Task 1
```{r}
houses <- read.csv("data/ames.csv") %>% 
  clean_names()
```

# first looking back at tests for normality
# try for sales_price data we used earlier

```{r}
library(ggpubr)
ggdensity(houses$sale_price, 
          main = "Density plot of sales_price",
          xlab = "sales price")
```
```{r}
ggqqplot(houses$sale_price)
```

```{r}
sw_test <- shapiro.test(houses$sale_price)
sw_test
#is p value >0.05?
sw_test$p.value >0.05
#false ! so the p value is less than 0.05 suggesting that the data are significantly different from a normal distribution.
```

```{r}
  qqnorm(houses$sale_price)
  qqline(houses$sale_price)
```
```{r}
houses %>% 
  select(lot_area) %>% 
  summarise(skewness = skewness(lot_area, type = 1))
```

#Task 2 - taking a look at the lot_area data

```{r}
library(ggpubr)
ggdensity(houses$lot_area, 
          main = "Density plot of lot area",
          xlab = "lot area")
```
```{r}
ggqqplot(houses$lot_area)
```

```{r}
sw_test <- shapiro.test(houses$lot_area)
sw_test
#is p value >0.05?
sw_test$p.value >0.05
#no. so the p value is less than 0.05 suggesting that the data are significantly different from a normal distribution.
```

```{r}
  qqnorm(houses$lot_area)
  qqline(houses$lot_area)
```


```{r}
houses %>% 
  ggplot()+
  aes(x = lot_area) +
  geom_histogram(bins=100)
```
```{r}
area_stats <- houses %>% 
  summarise(min = min(lot_area),
            mean = mean(lot_area),
            max = max(lot_area),
            sd = sd(lot_area),
            IQR = as.numeric(quantile(lot_area,0.75)-quantile(lot_area,0.25))
            )
```

# The lot area distribution is heavily right skewed 



# Now for some re-sampling - try some different methods

# first investigate data by taking a repeated (1000 times) sample of size 200
```{r}
sample_200 <- houses %>%
  select(lot_area) %>% 
  rep_sample_n(size = 200, reps = 1000)

sample_200 %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram()
  
summary_sample_200 <- sample_200 %>%
  summarise(
    mean_lot_area = mean(lot_area)) 

summary_sample_200 %>% 
  ggplot() +
  aes(x =  mean_lot_area) +
  geom_histogram()
```
This plot show sthe distribution of estimates of the mean from samples of the data.

The bootstrap method, repeatedly samples the data with replicates. So we create 1000(say) different versions of the dataset using sampling with replacement

```{r}
  houses %>% 
    select(lot_area) %>% 
    rep_sample_n(size = nrow(houses), replace = TRUE, reps = 1000) %>%
    ungroup() %>% 
    filter(replicate < 10) %>% 
    ggplot((aes(x = lot_area))) +
    geom_histogram(bins=100) +
    facet_wrap(~replicate)
```
```{r}
bootstrap_resample <- houses %>% 
    select(lot_area) %>% 
    rep_sample_n(size = nrow(houses), replace = TRUE, reps = 5000) %>%
    #ungroup() %>% 
    summarise(mean_lot_area_boot = mean(lot_area)
              )
```
now we have 5000 estimates of the mean

```{r}
bootstrap_resample %>% 
  ggplot +
  aes(x=mean_lot_area_boot)+
  geom_histogram()
```

```{r}
ci_95 <- bootstrap_resample %>% 
            summarise(mean = mean(mean_lot_area_boot),
            lower_bound = quantile(mean_lot_area_boot, probs = 0.025),
            upper_bound = quantile(mean_lot_area_boot, probs = 0.975)
            )
```

```{r}
bootstrap_resample %>% 
  ggplot +
  aes(x=mean_lot_area_boot)+
  geom_histogram(bins=30) +
  geom_vline(xintercept = ci_95$mean, col = "blue") +
  geom_vline(xintercept = ci_95$lower_bound, col = "red") +
  geom_vline(xintercept = ci_95$upper_bound, col = "red")
```

# do the same thing as above but use the infer method

#Task 3
```{r}
# specify that we want to look at lot_area variable, 5000 reps, and want to calculate the mean
infer_resample <- houses %>%
  specify(response = lot_area) %>%
  generate(reps = 5000, type = "bootstrap") %>%
  calculate(stat = "mean")
```

#Task 4
```{r}
infer_ci_95 <- infer_resample %>%
  get_confidence_interval(level = 0.95, type = "percentile")
infer_ci_95
```
```{r}
infer_resample %>%
  visualise(bins=30) +
  shade_confidence_interval(endpoints = infer_ci_95)
```

#Task 5 
```{r}
infer_ci_99 <- infer_resample %>%
  get_confidence_interval(level = 0.99, type = "percentile")
infer_ci_99
```

```{r}
infer_ci_95$upper_ci - infer_ci_95$lower_ci
```

```{r}
infer_ci_99$upper_ci - infer_ci_99$lower_ci
```

The 99% CI is a wider distribution - to have more confidence in our estimate we need it to encompass a wider range of the data. 

#Task 6
```{r}
mean_infer <- infer_resample %>% 
  summarise(mean = mean(stat))

mean_infer
```

# Extension
# Run the same sequence for proportion off houses built before 1920.

```{r}
houses <- houses %>% 
  mutate(pre_1920 = ifelse(year_built<1920,"yes","no")) 
```
```{r}
houses %>% 
  ggplot(aes(x=pre_1920)) +
  geom_bar() 
```

```{r}
houses %>%     
  summarise(prop_pre_1920 = sum(pre_1920)/n()) 
```


```{r}
samples = c(200,400,600,800,1000,
            2000,3000,4000,5000,6000,7000,8000,9000,
            10000,15000,20000,25000,30000,40000,50000)
results=rep(0,length(samples))
analysis=tibble(samples,
                results_prop=results,
                results_ci_upper=results,
                 results_ci_lower=results)
```

# loop through sample sizes
```{r}
for (isamp in 1:length(analysis$samples)){

# specify that we want to look at lot_area variable, start with 200 reps
infer_resample <- houses %>%
  specify(response = pre_1920, success = "yes") %>%
  generate(reps = analysis$samples[isamp], type = "bootstrap") %>%
  calculate(stat = "prop")

# calc ci 95
infer_ci_95 <- infer_resample %>%
  get_confidence_interval(level = 0.95, type = "percentile")

analysis$results_ci_upper[isamp] <- infer_ci_95 %>%
  select(upper_ci) %>% 
  pull()
analysis$results_ci_lower[isamp] <- infer_ci_95 %>%
  select(lower_ci) %>% 
  pull()

#visualise
infer_resample %>%
  visualise(bins=30) +
  shade_confidence_interval(endpoints = infer_ci_95)

#capture estimate
prop_infer <- infer_resample %>% 
  summarise(mean = mean(stat))

analysis$results_prop[isamp] <- prop_infer %>% 
  pull()
}
```

```{r}
analysis %>% 
#visualise effect or sampling
ggplot() +
aes(x=samples, y=results_prop) +
geom_point() +
geom_errorbar(aes(ymin=results_ci_lower,ymax=results_ci_upper))
```

```{r}
analysis %>% 
#visualise effect or sampling
ggplot() +
aes(x=samples, y=results_prop) +
geom_line()
```


