---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(skimr)
```
```{r}
movies_data <- read_csv("data/movies.csv")
```
```{r}
head(movies_data,10)
```
noticed there is a column for date and this is being read in as a character string so may need converting if we want to analyse by date

```{r}
skim(movies_data)
```
shows there are lots of missing data in homepage and tagline
and a few in overview and release date and runtime

check column names - they all seem OK

```{r}
names(movies_data)
```
notes on how to check differences. filter for when they are not the same
original_title seems to be linked to original_language
```{r}
movies_data %>% 
  select(title, original_title, original_language) %>% 
  filter(title != original_title) %>% 
  arrange(title)  
```



Q2
```{r}
movies_selected <- movies_data %>% 
  select(original_title, runtime, budget)
movies_selected
```

Q3

```{r}
movies_selected %>% 
  summarise(title_na = sum(is.na(original_title)),
            runtime_na = sum(is.na(runtime)),
            budget_na = sum(is.na(budget))
            )
```
# summary(movies_selected) gies the same answer
# map(movies_selected, ~sum(is.na(.)))
# or use the lambda function 

```{r}
movies_selected %>% 
  summarise(across(everything(), ~sum(is.na(.x))))
```

there are just 2 missing values in runtime

Q4

removing cases where runtime=0
first checking it works
```{r}
movies_selected %>%
  mutate(runtime = na_if(runtime,0)) %>% 
  arrange(runtime) 
```
```{r}
movies_selected <- movies_selected %>%
  select(original_title,runtime,budget) %>% 
  mutate(runtime = na_if(runtime,0))
```

```{r}
summary(movies_selected["runtime"])
```

answer - 37 missing values

Q5
```{r}
movies_imputed <- movies_selected %>%
  mutate(runtime= coalesce(runtime, mean(runtime, na.rm =TRUE))) 
```

```{r}
summary(movies_data["runtime"])
```

```{r}
summary(movies_selected["runtime"])
```

```{r}
summary(movies_imputed["runtime"])
```
just checking impact of my corrections. Makes sense as runtime now ranges from 14 to 338 with no NA's       

Q6
```{r}
movies_imputed %>% 
  arrange(runtime) %>% 
  head(10)
```
```{r}
movies_imputed %>% 
  arrange(desc(runtime)) %>% 
  head(10)
```
 
# answer using slice
slice is subsetting data  by index - good for decimating, selecting all odd. all even, for example

```{r}
movies_selected %>% 
slice_max(runtime, n=10)
```
```{r}
movies_selected %>% 
slice_min(runtime, n=10)
```
       
       
```{r}
movies_imputed <- movies_imputed %>% 
  mutate(budget = if_else(budget<100,median(budget), budget)) %>% 
  arrange(budget)
```
       
check results of above
```{r}
summary(movies_imputed)
```

## Extension

Q1
```{r}
movies_budgets <- movies_imputed %>% 
  mutate(budget_type = case_when(
         budget >= 40e6 ~ "Big budget",
         budget >= 12e6 ~ "Medium budget",
         TRUE ~ "Small budget"
  )) %>% 
  arrange(desc(budget))
```
double check results of above
```{r}
movies_budgets %>% 
  group_by(budget_type) %>% 
  summarise(count=n(), min_budget=min(budget), max_budget=max(budget))
```
Q2

```{r}
my_isna_fun <- function(x) {
  sum(is.na(x))
}
```


```{r}
movies_data %>%
    select(where(is.numeric)) %>% 
    summarise(
    across(.cols = where(is.numeric),
                .fns = my_isna_fun)
                )
```

```{r}
movies_data %>%
    select(where(is.character)) %>% 
    summarise(
    across(.cols = where(is.character),
                .fns = my_isna_fun)
                )
```

```{r}
movies_data %>%
    select(where(is.numeric)) %>% 
    summarise(
    across(.cols = where(is.numeric),
                .fns = ~sum(is.na(.x)))
                )
```

```{r}
movies_data %>%
    select(where(is.character)) %>% 
    summarise(
    across(.cols = where(is.character),
                .fns = ~sum(is.na(.x)))
                )
```