---
title: "Probability Homework"
output: html_notebook
---

```{r}
library(gtools)
library(tidyverse)
source('prob.R')
```
#Recap

probability is the numerical measure of the liklihood of an event occuring.
It ranges between 0 (impossible) and 1 (certain) 

sample space = collection of all possible outcomes

event space = a subset of the sample space, the events we are interested in 


```{r}
titanic_data = read_csv("data/titanic.csv")
```

```{r}
titanic_data %>% 
  distinct(class)
```

```{r}
titanic_data <- titanic_data %>% 
  mutate(status = ifelse(class %in% c("1st","2nd","3rd"),"passenger","crew"))  
```
 
```{r}
#just check mutate worked
titanic_data %>% 
  filter(status == "crew") %>% 
  distinct(class)
```
 


## Calculate conditional probability of survived | crew and survived | passenger

```{r}
prob_survived_given_crew <- titanic_data %>% 
  filter(status == "crew") %>% 
  summarise(prob = sum(survived == "yes") /n())
```

```{r}
prob_survived_given_passenger <- titanic_data %>% 
  filter(status == "passenger") %>% 
  summarise(prob = sum(survived == "yes") /n())
```

```{r}
prob_survived_given_crew
prob_survived_given_passenger
```
The probability of surviving was higher if you were a passenger (0.38) than if you were a crew member (0.24)

# Probability questions
# Q1 - what is the probability of getting a number less than 5 when rolling a 6 sided dice

number of possible outcomes (sample_space) = 6  
event = {1,2,3,4}
prob is 4/6 or 2/3

#Q2 - how many possible outcomes when we throw three dice
each dice has 6 possible outcomes so possible outcomes from 3 dices is 6*6*6

```{r}
roll_3_dice <- rolldie(3)
nrow(roll_3_dice) # sample space is 216
```
 #Q3 - in an experiment with 4 dice
 sample space = total possible outcomes = 6^4 = 1296
 
```{r}
roll_4_dice <- rolldie(4) %>% 
  mutate(die_tot = rowSums(.))
```

```{r}
sum_bet_5_7 <- roll_4_dice %>% 
  filter(die_tot %in% c(5,6,7))
sum_bet_5_7
```
```{r}
prob = nrow(sum_bet_5_7)/nrow(roll_4_dice)
prob
```
# In answering part 2 we are assuming that order is not important?

#Q4 
Five chairs placed in a row, 5 people with different heights
people sit down randomly - what is the possibility of they sit in decreasing height L to R
peope are unique - we cant have repeats
sample space is 5! (% factorial) or 5*4*3*2*1 = 120
event space is 1 (only 1 way to sit in this specific order)
prob = 1/120

```{r}
# create vector of choices contains 5 people of heights H1 to H5
#H5 is the tallest and H1 is the smallest. n = 5
person_height <- c("H1", "H2", "H3", "H4", "H5")

# sit 5 at a time, r = 5
# repeats.allowed = FALSE - only one person per chair
chair_permutations <- as_tibble(permutations(n = 5, r = 5, v = person_height, repeats.allowed = FALSE)) 

prob = 1/nrow(chair_permutations)
prob
```

#Q5 I roll six dice: what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
dice are not unique so sample space is 6^6
```{r}
1/6^6
```
#Q5 - what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?

```{r}
# calculate sample space (6^6)
roll_6_dice <- rolldie(6) 
roll_6_dice
```

```{r}
# alternatively calculate event space (6!)
dice_permutations <- as_tibble(permutations(n = 6, r = 6, v = (c(1:6)), repeats.allowed = FALSE)) 
dice_permutations
```

```{r}
prob <- factorial(6)/6^6
prob
prob <- nrow(dice_permutations)/nrow(roll_6_dice)
prob
```

```{r}
# could also use isin function
rolldie(6) %>% 
 mutate(check = isin(., c(6,5,4,3,2,1), ordered = FALSE)) %>% 
 filter(check) %>% 
 summarise(prob = n()/nrow(roll_6_dice))  
```

#Q6 If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

card deck has 52 cards
number of black cards is half of that 26 so sample space s 26
number of spades is half of that so prob is 13/26? 0.5

```{r}
cards(jokers = FALSE) %>%
 mutate(black = suit %in% c("Spade","Club")) %>%   
 filter(black) %>% 
 filter(suit == "Spade")  
```

