---
title: "Homework"
output: html_notebook
---

```{r}
library(tidyverse)
library(CodeClanData)
library(skimr)
library(GGally) # library for creating those insane correllelograms :)
```

```{r}
glimpse(backpack)
names(backpack)
skim(backpack)
```

```{r}
backpack_data <- backpack
```

```{r}
backpack_data %>% 
ggplot((aes( x= backpack_weight))) +
 geom_histogram(binwidth=1, boundary = 0.5, colour = "white")
```
We can say the majority (12) of the students are carrying a bag weighing 10kg. A small number of students are carryng a very heavy bags, including one extremely heavy bag of 35kg. 


```{r}
backpack_data %>% 
ggplot((aes( x= backpack_weight))) +
 geom_histogram(binwidth=5, boundary = 0.5, colour = "white")
```

The majority of students are carrying a bag between 6-10 kg

```{r}
backpack_data %>% 
ggplot((aes(x = sex, y = backpack_weight))) +
 geom_boxplot() +
 ylim(0, max(backpack_data$backpack_weight))
```

```{r}
ggplot(backpack_data) + 
  aes(backpack_weight, fill = sex) + 
  geom_bar(position = position_dodge2(width = 1, preserve = "single"))
  # preserve = single fixes the situations where bar widths are not all the same size
```


```{r}
backpack_data %>% 
ggplot((aes( x= backpack_weight, fill = sex))) +
 geom_histogram(binwidth= 1, boundary = 0.5, colour = "white")
```

Male students on average, had slightly heavier bags. The range of backpack weights carried by male and female students was similar.

```{r}
backpack_data %>% 
ggplot(aes(x = body_weight, y = backpack_weight)) +
 geom_point() +
 geom_smooth(method = "lm")  
```

```{r}
backpack_data %>% 
ggplot(aes(x = body_weight, y = backpack_weight)) +
 geom_point() +
 geom_smooth(method = "loess", span = 1)  
```


```{r}
backpack_data %>% 
ggplot(aes(x = body_weight, y = backpack_weight, colour = sex)) +
 geom_point() +
 geom_smooth(method = "loess", span = 1)  
```
There could be a number of reasons why backpack and body weight might vary. A larger person has larger clothes and therefore their backpack weight would be positively correlated with body weight. In most able-bodied people, strength could also be related to bodyweight, and in a group scenario stronger people may be required to carry more. At very high bodyweights fitness may have an impact and so the largest people may adjust by choosing to carry a lighter backpack.

From the data, there does not appear to be a linear relationship between body weight and backpack weight. Fitting a more complex smoother could indicate that backpack weight initially increases with body weight but then decreases again. A similar pattern is observed when fitting to male and female groups separately. 

However the confidence bands on both the linear fit and loess smoother are wide and many points lie outside of these bands. I dont think it's possible to have any confidence in these fitted trends. 


```{r}
backpack_data %>%  
  select(back_problems) %>% 
  group_by(back_problems) %>% 
  summarise(count = n(), prop=100*(n()/nrow(backpack_data)), tot = nrow(backpack_data))
```


```{r}
group_by(backpack_data,sex) %>% 
          summarise(count = n())
```

```{r}
backpack_data %>%  
  select(back_problems, sex) %>% 
  group_by(sex, back_problems) %>% 
  summarise(count = n(), prop=100*(n()/nrow(backpack_data)), tot = nrow(backpack_data)) %>% 
  arrange(desc(sex))
```
```{r}
backpack_data <-  rowid_to_column(backpack_data)

```


```{r}
backpack_data %>% 
mutate(back_pain = ifelse(back_problems == 1, 
                          "back_pain", "no_back_pain")) %>% 
ggplot(aes(x = sex, fill = back_pain)) +
geom_bar(position = "fill", colour = "black") +
labs(y = "% with/without backpain", 
     x = "Sex",
     title = "Occurrence of back pain in male and female students") +
scale_y_continuous(breaks=c(seq(0, 1, by = 0.1)),labels = scales::percent)
```



There are 100 records! So 32% of the students had back problems. There were proportionally more of the female students(31/55) with back problems than male students (8/45)

```{r}
select(backpack_data, backpack_weight, ratio, body_weight, sex, back_problems) %>% 
ggpairs()
```

```{r}
backpack_data %>% 
ggplot(aes(x = body_weight, y = backpack_weight, colour = sex)) +
 geom_point() +
 facet_grid(~back_problems)
```

```{r}
backpack_data %>% 
ggplot(aes(x = body_weight, y = backpack_weight)) +
 geom_point() +
 geom_smooth(method="lm") +
 facet_grid(sex ~ back_problems)
```
Im not sure I can see any meaningful relationship here? Maybe for males with backpain, there is an inverse relationship between backpack_weight and bodyweight. For higher body_weights there is a tendency for backpack_weight to be lower? 

```{r}
backpack_data %>% 
  ggplot(aes(x = as.factor(back_problems), y = ratio)) +
  geom_boxplot() +
  facet_grid(~sex)
```
Looking at it like this it seems that those people with back problems had a higher overall body_weight:backpack_weight ratio. This was true of both sexes.


```{r}
backpack_data %>%
filter(year >0) %>%   
mutate(back_pain = ifelse(back_problems == 1, 
                          "back_pain", "no_back_pain")) %>% 
ggplot(aes(x = year, fill = back_pain)) +
geom_bar(position = "stack" , colour = "black") +
labs(y = "# with/without backpain", 
     x = "Year",
     title = "Back pain by year group") +
scale_x_continuous(breaks=c(1:6))
```


```{r}
backpack_data %>%
filter(year >0) %>%   
mutate(back_pain = ifelse(back_problems == 1, 
                          "back_pain", "no_back_pain")) %>% 
ggplot(aes(x = year, fill = back_pain)) +
geom_bar(position = "fill", colour = "black") +
labs(y = "% with/without backpain", 
     x = "Year",
     title = "Back pain by year group") +
scale_x_continuous(breaks=c(1:6)) +
scale_y_continuous(breaks=c(seq(0, 1, by = 0.1)),labels = scales::percent)
```
Year 6 had the lowest proportion of back pain amongst students with 0 incidents of back_problems. Year 3 had the highest with 


```{r}
backpack_data %>%
filter(year > 0) %>%   
ggplot(aes(x = as.factor(year), y = backpack_weight)) +
geom_boxplot() +
labs(y = "back_pack weight (kg?)", 
     x = "Year of Study",
     title = "Back_pack weight by year group") 
```
Back pack weight by year group? Possibly an increase in average backpack weight from year 1 - year 5 and then a decrease?

```{r}
backpack_data %>%
filter(year > 0) %>%   
ggplot(aes(x = as.factor(year), y = backpack_weight)) +
geom_point() +
labs(y = "back_pack weight (kg?)", 
     x = "Year of Study",
     title = "Back_pack weight by year group") 
```
Looking at it like this makes me think that the sample is small and the patterns seen in the boxplot may not be significant.
